---
title: "Creating the ``r params$package_name`` R package"
author: "Jacob Bien"
date: "Sept 8, 2022"
knit: litr::render
params:
  package_name: "rleastsquares" # <-- change this to your package name
  package_parent_dir: "." # <-- relative to this file's location
---

<!-- This Rmd file contains all the code needed to define an R package.  Press "Knit" in RStudio or more generally run `rmarkdown::render("name-of-this-file.Rmd")` to generate the R package.  Remember that when you want to modify anything about the R package, you should modify this document rather than the package that is outputted.
-->

\newcommand{\real}{\mathbb R}

Note: This package was coded in a live coding demo.  

## Package setup

We start by specifying the information needed in the DESCRIPTION file of the R package.

```{r package-setup, message=FALSE, results='hide'}
usethis::create_package(
  path = ".",
  fields = list(
    Package = params$package_name,
    Version = "0.0.0.9000",
    Title = "A Package That Does Least Squares",
    Description = "This package says does least squares.",
    `Authors@R` = person(
      given = "Jacob",
      family = "Bien",
      email = "jbien@usc.edu",
      role = c("aut", "cre")
      )
  )
)
usethis::use_mit_license(copyright_holder = "J. Bien")
```

Suppose we have a response vector $y\in\real^n$ and a data matrix $X\in\real^{n\times p}$.

We want to find the solution to the problem

$$
\min_{\beta\in\real^p}\|y-X\beta\|^2
$$

We'll assume that $X$ is full rank with $n > p$.  We know that the solution is given by

$$
\hat\beta=(X^TX)^{-1}X^Ty.
$$
Our main function does exactly that!

```{r}
#' Get the OLS solution
#' 
#' @param y our response, which is an n-vector
#' @param X our data matrix, which is n by p
#' @export
do_least_squares <- function(y, X) {
  if(nrow(X) != length(y))
    stop("The number of rows of X must match the length of y.")
  as.numeric(solve(crossprod(X), crossprod(X, y)))
}
```

Let's see how this works!

```{r}
set.seed(123)
n <- 100
p <- 1
x <- cbind(1, matrix(rnorm(n*p), n, p))
beta_star <- c(2, 0.5)
sigma <- 0.1
y <- x %*% beta_star + sigma * rnorm(n)
```


```{r}
betahat <- do_least_squares(y, x)
```

```{r}
plot(x[, 2], y)
abline(betahat[1], betahat[2], col = 2, lwd=2)
```

Let's see how this compares to `lm`'s answer.


```{r}
fit_lm <- lm(y ~ x[, 2])
```

```{r}
fit_lm$coefficients
```

Compare that to... 

```{r}
betahat
```


```{r}
testthat::test_that("do_least_squares() works", {
  set.seed(123)
  n <- 100
  p <- 1
  x <- cbind(1, matrix(rnorm(n*p), n, p))
  beta_star <- c(2, 0.5)
  sigma <- 0.1
  y <- x %*% beta_star + sigma * rnorm(n)
  fit_lm <- lm(y ~ x[, 2])
  # do lm and our function give the same coefficient vector?
  testthat::expect_equal(do_least_squares(y, x),
                         as.numeric(fit_lm$coefficients))
  # do we get the desired error when there is a length mismatch?
  testthat::expect_error(do_least_squares(y[-1], x), "must match")
})
```

Code chunks that have one or more lines starting with `test_that(` (or `testthat::test_that(`) are added to the package as tests.

## Documenting the package and building

We finish by running commands that will document, build, and install the package.  It may also be a good idea to check the package from within this file.

```{r}
rm(list = ls())
litr::document() # <-- use instead of devtools::document()
# devtools::build()
# devtools::install()
# devtools::check(document = FALSE)
```


